---
title: "Bayesian Logistic Regression on Media Spending Data"
author: "Taiwo Owoseni"
date: "23/12/2022"
output: pdf_document
---

## Introduction

This analysis aims to use a Bayesian approach to build a logistic regression model that classifies the various campaigns overtime at the Division level. The data used the analysis was sourced from [Kaggle](https://www.kaggle.com/datasets/yugagrawal95/sample-media-spends-data). Visit to link to learn more about the data\

## Import Libraries

```{r library, include=FALSE}
library(bayesrules)
library(tidyverse)
library(broom)
library(rstan)
library(bayesplot)
#library(broom.mixed)
library(cowplot)
library(janitor)

```

## About the Data

```{r load data}
data <- read.csv("Sample Media Spend Data.csv")
head(data)

```

```{r row and column count}
paste( "Number of observations:",nrow(data))
paste( "Number of observations:",ncol(data))

```

The data set contains 3051 observations and ten variables.

## Statistical Inquiries

Suppose we are interested in the following statistical inquiries:

1.  How is continuous `Google_Impressions` associated with `Division`?

2.  Is there a numerical result that can quantify this association?

We are interested in the following variables:

-   `Division`: Subjects division level. The various campaigns overtime at the Division level.. It is a categorical variable with 26 levels `A-Z` but for this study, we will focus on two levels `A and Z`

-   `Google_Impressions` : The number of times content entered a subject's screen while on google. It's a continuous variable 

For this study, `Division` is the `dependent` variable and `Google_Impressions` is the `independen`t variable.

```{r filter data}
nw.data <- data %>%
  select(Google_Impressions, Division) |>
  filter(Division %in%  c("A", "Z"))
  
levels(nw.data$Division)
nw.data

```

### Scaling the Data

```{r range of impression}
paste('The lowest impression in the data is: ', min(nw.data$Google_Impressions))
paste('The highest timpresison in the data 1s: ', max(nw.data$Google_Impressions))
```

The `Google impressions` in the data across all divisions range from 7 to 17,150,439. For this reason, the log transformation of the `Google_impression` will be used in this analysis.

```{r log transform }
nw.data$Log_Google_Impressions <- log(nw.data$Google_Impressions)

paste('The lowest impression in the data is: ', min(nw.data$Log_Google_Impressions))
paste('The highest timpresison in the data 1s: ', max(nw.data$Log_Google_Impression))
```

Due to the log transformation, the range of the scaled google impression is smaller when compared to the previous values.

```{r}
nw.data <- nw.data %>%
  mutate(Division = case_when(
    Division == "A" ~ 1,
    Division == "Z" ~ 0
  )) |> select(c('Log_Google_Impressions', 'Division'))
head(nw.data)
```
