---
title: Analyze Causality - Causal influence of Self-rated enjoyment of MDS program on expected salary upon graduation 
author: Taiwo Owoseni
output:
  pdf_document:
    toc: true
    toc_depth: 2
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(tools)
library(scales)
library(broom)
library(MASS)
library(rmarkdown)
```

### Problem Statement

Using a survey data form 2019 to 2020 of past MDS students

1.  Analyse whether a person's self-rated enjoyment of the MDS program
    ($X$) had any causal influence on their expected salary upon
    graduation ($Y$)?
2.  Is there a statistical difference between the data of both years?
3.  Does the model without confounding variables perform better than a
    model with confounding variables

The following are the features in the data set set and their
description:

+---------------+----------------------------------------------+
| Features      | Description                                  |
+===============+==============================================+
| **Salary      | Previous salary could be confounding as it   |
| before MDS**  | will affect future salary expectations and   |
|               | affect happiness in the program. For         |
|               | example, a low prior salary may make a       |
|               | person happier to enter MDS and have a       |
|               | potentially higher future salary.            |
+---------------+----------------------------------------------+
| **Years of    | Years of work experience could be            |
| professional  | confounding as it will affect future salary  |
| work          | expectations. A person with more years of    |
| experience**  | work experience might not be as happy to     |
|               | enter MDS.                                   |
+---------------+----------------------------------------------+
| **Confidence  | Confidence in data science skills before     |
| in data       | starting MDS would likely affect future      |
| science       | salary expectations and directly influence   |
| skills before | how happy they were in the program. If a     |
| starting      | person is more confident, then they will     |
| MDS**         | likely have an easier time and be happier.   |
+---------------+----------------------------------------------+
| **Frequency   | Doing optional questions in labs indicates a |
| of completing | person is likely finishing all of the        |
| optional      | required content with relative ease or likes |
| questions in  | a challenge. If this is the case, they could |
| labs**        | be happier with MDS and expect a larger      |
|               | salary to reflect their skill set.           |
+---------------+----------------------------------------------+
| **Actively    | If people are actively applying for jobs in  |
| applying for  | block 6 of MDS, it could be affirming that   |
| jobs in block | MDS has prepared them adequately or the      |
| 6 of MDS**    | opposite. This would affect their feelings   |
|               | about the program and, consequently, salary  |
|               | expectations.                                |
+---------------+----------------------------------------------+
| **Current     | A baseline for happiness. If they are not    |
| happiness     | happy, they are not happy with MDS, and then |
| level**       | maybe they do not expect much of a salary.   |
+---------------+----------------------------------------------+
| **Frequency   | Attending career events could increase       |
| of attending  | Confidence in job opportunities and,         |
| MDS career    | subsequently, salary. If the events are      |
| events**      | useful, then the feelings about MDS would    |
|               | likely change for the better.                |
+---------------+----------------------------------------------+

> **Note:** The columns names in the data sets from
> *`import_id_qid172807697 to import_id_qid93`are renamed as and*
> represent the columns *`salary_exp_post_grad` (***Salary before MDS)**
> *to `freq_attend_mds_career_events`* **(Frequency of attending MDS
> career events)**

> **Note:** Given that the problem statement can only be realistically
> answered with an observational study, confounders were considered when
> collecting the data (via the planned survey) and carrying out the
> analysis

### Exploring Data sets and cleaning

**STEP 1:** Read data set as csv from local folder, skipped the first
two rows.

**STEP 2:** Clean column names so they are unique and consist only of
the **`_`** character, numbers, and letters.

**STEP 3:** Print out first 5 rows in `salary_df_2019`

```{r message=FALSE}
salary_df_2019 <- read_csv("data/salary_survey_responses_2019.csv", skip = 2) %>%
  clean_names()

salary_df_2020 <- read_csv("data/salary_survey_responses_2020.csv", skip = 2) %>%
  clean_names()

head(salary_df_2019)
```

The data was collected **in 2019 and 2020** via a survey with 55 and 44
respondents each.

```{r}
paste('salary_df_2019 has' , nrow(salary_df_2019), 'rows')
paste('salary_df_2020 has' , nrow(salary_df_2020), 'rows')

```

#### Data Wrangling and Merging data sets

**STEP 1**: Renamed data set using the description given in Table 1 as
follows

-   `import_id_qid172807697` as `salary_exp_post_grad`
-   `import_id_qid172807701_1` as `mds_self_rated_enjoy`
-   `import_id_qid96` as `salary_pre_mds`
-   `import_id_qid172807686` as `work_exp`
-   `import_id_qid98_1` as `ds_skill_confidence`
-   `import_id_qid172807685` as `does_optional_qs`
-   `import_id_qid92` as `currently_job_searching`
-   `import_id_qid99_1` as `baseline_happiness`
-   `import_id_qid93` as `freq_attend_mds_career_events`

**STEP 2:** Create a column called `year` with the corresponding labels:
`2019` or `2020`.

**STEP 3:** Drop all observations with missing data.

**STEP 4:** Merge `salary_df_2019` and `salary_df_2020` into a single
data frame called `salary_df`.

**STEP 5:** Print out first 5 rows in `salary_df`

```{r data wrangling solution 1}
# BEGIN SOLUTION

salary_df_2019 <- salary_df_2019 %>%
  dplyr::select(contains("qid")) %>%
  rename(
    salary_exp_post_grad = import_id_qid172807697,
    mds_self_rated_enjoy = import_id_qid172807701_1,
    salary_pre_mds = import_id_qid96,
    work_exp = import_id_qid172807686,
    ds_skill_confidence = import_id_qid98_1,
    does_optional_qs = import_id_qid172807685,
    currently_job_searching = import_id_qid92,
    baseline_happiness = import_id_qid99_1,
    freq_attend_mds_career_events = import_id_qid93
  ) %>%
  drop_na() %>%
  mutate(year = 2019)

salary_df_2020 <- salary_df_2020 %>%
  dplyr::select(contains("qid")) %>%
  drop_na() %>%
  mutate(year = 2020)
colnames(salary_df_2020) <- colnames(salary_df_2019)

salary_df <- bind_rows(salary_df_2019, salary_df_2020)

head(salary_df)
```

**STEP 1**: Get all the unique values for all columns in `salary_df`

```{r function_unique_values}

unique_values<- function(df){
  sapply(df, unique)
}
unique_values(salary_df)
```

#### Column Renaming and Value Factoring

**STEP 1:** Use `toTitleCase()` to change the level names of
`freq_attend_mds_career_events` to *Title Style*.

**STEP 2:** Change the factor level `0 - 1 Years` to `Less than 1 Year`
in `work_exp`.

**STEP 3:** Convert columns `does_optional_qs`,
`currently_job_searching`, and `year` to **NOMINAL factor-type**.

**STEP 4:** Convert the rest of the columns to **ORDERED factor-type**.

**STEP 5**: Make sure that factors `salary_exp_post_grad`,
`mds_self_rated_enjoy`, `salary_pre_mds`, `work_exp`,
`ds_skill_confidence`, `baseline_happiness`, and
`freq_attend_mds_career_events` have the correct level order from left
to right via function `levels()`. If not, reorder these levels according
to the order detailed at the beginning of this exercise.

```{r data wrangling}
# BEGIN SOLUTION

salary_df <- salary_df %>%
  mutate(
    freq_attend_mds_career_events =
      toTitleCase(freq_attend_mds_career_events)
  ) %>%
  mutate(work_exp = case_when(
    work_exp == "0 - 1 Years" ~ "Less than 1 Year",
    TRUE ~ work_exp
  )) %>%
  mutate(
    does_optional_qs = factor(does_optional_qs),
    currently_job_searching = factor(currently_job_searching),
    year = factor(year),
    salary_exp_post_grad = factor(salary_exp_post_grad, ordered = TRUE),
    mds_self_rated_enjoy = factor(mds_self_rated_enjoy, ordered = TRUE),
    salary_pre_mds = factor(salary_pre_mds, ordered = TRUE),
    work_exp = factor(work_exp, ordered = TRUE),
    ds_skill_confidence = factor(ds_skill_confidence, ordered = TRUE),
    baseline_happiness = factor(baseline_happiness, ordered = TRUE),
    freq_attend_mds_career_events = factor(freq_attend_mds_career_events,
      ordered = TRUE
    )
  ) %>%
  mutate(
    salary_exp_post_grad = fct_relevel(
      salary_exp_post_grad,
      "Less than $60,000",
      "$60,000 to $80,000",
      "$80,001 to $100,000",
      "$100,001 to $120,000",
      "More than $120,000"
    ),
    salary_pre_mds = fct_relevel(
      salary_pre_mds,
      "Less than $60,000",
      "$60,000 to $80,000",
      "$80,001 to $100,000",
      "$100,001 to $120,000",
      "More than $120,000"
    ),
    work_exp = fct_relevel(
      work_exp,
      "Less than 1 Year",
      "1 - 4 Years",
      "4 - 7 Years",
      "7 - 10 Years",
      "10+ Years"
    ),
    freq_attend_mds_career_events = fct_relevel(
      freq_attend_mds_career_events,
      "Not Often",
      "Sometimes",
      "Often"
    )
  )

# END SOLUTION

```

**Previous salary prior to MDS (`salary_pre_mds`) has the following 5
levels:**

-   `Less than $60,000`
-   `$60,000 to $80,000`
-   `$80,001 to $100,000`
-   `$100,001 to $120,000`
-   `More than $120,000`

**Years of professional work experience prior to MDS (`work_exp`) has
the following 5 levels:**

-   `0 - 1 Years`
-   `1 - 4 Years`
-   `4 - 7 Years`
-   `7 - 10 Years`
-   `10+ Years`

**Confidence in data science skill when first starting MDS on a scale of
`1` - `4`. With `4` being very confident and `1` being not confident.
(`ds_skill_confidence`) has the following 4 levels:**

-   `1`
-   `2`
-   `3`
-   `4`

**Does optional questions in labs (`does_optional_qs`) has the following
2 levels:**

-   `Yes`
-   `No`

**Currently applying for data science jobs (`currently_job_searching`)
has the following 2 levels:**

-   `Yes`
-   `No`

**Current happiness level on a scale of `1`-`4`. With `4` being very
happy and `1` being not happy. (`baseline_happiness`) has the following
4 levels:**

-   `1`
-   `2`
-   `3`
-   `4`

**Frequently attending MDS career events
(`freq_attend_mds_career_events`) has the following 3 levels:**

-   `Not often`
-   `Sometimes`
-   `Often`

### Visualization

Make eight suitable plots of `salary_exp_post_grad` versus each one of
the rest of factor-type variables except `year`. Nonetheless, in these
eight plots, include panels per `year`.

> **Hint:** If you are using the same class of plot eight times, it
> would be practical to build a function first.

```{r plotting solution}
# BEGIN SOLUTION

# Plotting function

plotting_function <- function(dataset, response, exp_variable, x_axis, plot_title) {
  dataset.prop.summary <- dataset %>%
    group_by(year, eval(parse(text = exp_variable)), eval(parse(text = response))) %>%
    summarize(n = length(eval(parse(text = exp_variable)))) %>%
    mutate(prop = n / sum(n))

  colnames(dataset.prop.summary)[2:3] <- c(exp_variable, response)

  ggplot(dataset.prop.summary, aes(
    x = eval(parse(text = exp_variable)),
    y = prop, fill = eval(parse(text = response))
  )) +
    ggtitle(plot_title) +
    geom_bar(stat = "identity", width = 0.7, colour = "black", lwd = 0.1) +
    geom_text(aes(label = ifelse(prop >= 0.01,
      paste0(sprintf("%.0f", prop * 100), "%"), ""
    )),
    position = position_stack(vjust = 0.5), colour = "red3",
    fontface = "bold", size = 3
    ) +
    scale_y_continuous(labels = percent_format()) +
    labs(y = "Percent", x = x_axis, fill = "") +
    theme(
      axis.text.x = element_text(size = 11, angle = 0),
      axis.text.y = element_text(size = 11, angle = 0),
      axis.title = element_text(size = 12),
      legend.text = element_text(size = 9, margin = margin(r = 0.5, unit = "cm")),
      legend.title = element_text(size = 11, face = "bold")
    ) +
    guides(fill = guide_legend(title = "Salary\nExpectation\nAfter Graduation\n(CAD)")) +
    facet_grid(~year) +
    coord_flip()
}

```

#### Percentage of `salary_exp_post_grad` vs `mds_self_rated_enjoy`

Regarding the explanatory variable of interest `mds_self_rated_enjoy`,
there is a more diverse composition of salary expectation as the level
of happiness increases, it leans more towards higher salary
expectations.

```{r message=FALSE, fig.width=10.5}
# BEGIN SOLUTION
plotting_function(
  salary_df, "salary_exp_post_grad", "mds_self_rated_enjoy",
  "Self-Rated Enjoyment", "% Salary Expectation Post Graduation \n by Self Rated Enjoyment in 2019 and 2020 across Four(4) levels"
)

```

#### Percentage of `salary_exp_post_grad` vs `salary_pre_mds`

In general, regardless of the year, the respondents tend to expect a
higher salary once they graduate from MDS compared to their previous
salary.

```{r, message=FALSE, fig.width=10.5}
# BEGIN SOLUTION
plotting_function(
  salary_df, "salary_exp_post_grad", "salary_pre_mds",
  "Salary Before Starting MDS", "% Salary Expectation Post Graduation by\n Salary Before Starting MDS in 2019 and 2020 across Five (5) levels"
)

```

#### Percentage of `salary_exp_post_grad` vs `work_exp`

Graphically, the respondents with more years of experience tend to
expect higher salaries once they graduate from MDS. This is more
noticeable in 2019 than in 2020.

```{r, message=FALSE, fig.width=10.5}
# BEGIN SOLUTION
plotting_function(
  salary_df, "salary_exp_post_grad", "work_exp",
  "Years of Previous \n Work Experience",  "% Salary Expectation Post Graduation by Years of Previous Work Experience in 2019 and 2020 across Five (5) levels"
)

```

#### Percentage of `salary_exp_post_grad` vs `ds_skill_confidence`

An higher level of confidence before starting MDS is graphically more
associated with a higher salary expectation.

```{r, message=FALSE, fig.width=10.5}
# BEGIN SOLUTION
plotting_function(
  salary_df, "salary_exp_post_grad", "ds_skill_confidence",
  "Confidence in Data Science \n Skills Before Starting MDS", "% Salary Expectation Post Graduation by Confidence in Data Science Skills Before Starting MDS in 2019 and 2020 across Four (4) levels"
)

```

#### Percentage of `salary_exp_post_grad` vs `does_optional_qs`

There are no large graphical differences in the respondent shares
between both years. Also, those who perform optional questions also tend
to expect a higher salary

```{r, message=FALSE, fig.width=10.5}
# BEGIN SOLUTION
plotting_function(
  salary_df, "salary_exp_post_grad", "does_optional_qs",
  "Performs Optional Questions\n on the Labs",  "% Salary Expectation Post Graduation by Performs Optional Questions on the Labs in 2019 and 2020 across Two (2) levels"
)
```

#### Percentage of `salary_exp_post_grad` vs `currently_job_searching`

The plot does not provide a clear trend between the respondent shares in
both levels of the confounder. This also applies between both years.

```{r, message=FALSE, fig.width=10.5}
# BEGIN SOLUTION
plotting_function(
  salary_df, "salary_exp_post_grad", "currently_job_searching",
  "Currently Applying for \n Data Science Jobs", "% Salary Expectation Post Graduation by Currently Applying for Data Science Jobs in 2019 and 2020 across Two (2) levels"
)
# END SOLUTION
```

#### Percentage of `salary_exp_post_grad` vs `baseline_happiness`

It is pretty noticeable that a higher happiness level is also associated
with higher salary expectations. This trend applies for both years.

```{r, message=FALSE, fig.width=10.5}
# BEGIN SOLUTION
plotting_function(
  salary_df, "salary_exp_post_grad", "baseline_happiness",
  "Current Happiness Level", "% Salary Expectation Post Graduation by Current Happiness Level in 2019 and 2020 across Four (4) levels"
)
# END SOLUTION
```

#### Percentage of `salary_exp_post_grad` vs `freq_attend_mds_career_events`

Those who do not often attend MDS career events also expect higher
salaries after graduation.

```{r, message=FALSE, fig.width=10.5}
# BEGIN SOLUTION
plotting_function(
  salary_df, "salary_exp_post_grad", "freq_attend_mds_career_events",
  "Frequency of Attending \n MDS Career Events", "% Salary Expectation Post Graduation by Current Happiness Level in 2019 and 2020 across Three (3) levels"

)
# END SOLUTION
```

### Regression Model

Given that `salary_exp_post_grad` is an ordinal response, the most
suitable regression model is the ordinal logistic regression. To fit the
model with the package `MASS`, we use the function `polr()`, which
obtains the corresponding estimates.

Let us begin by just using the $X$ (`mds_self_rated_enjoy`) and $Y$
(`salary_exp_post_grad`) of interest in this observational study to
estimate a regression model with these two variables.

However, before starting with the model fitting, it is important to
highlight something regarding `mds_self_rated_enjoy` and the rest of the
**non-binary** survey questions . These variables are ordinal. Thus,
when fitting a regression with them as explanatory variables (regardless
of whether the regression is classical, binomial logistic, count-type,
etc.), we are likely interested in assessing the statistical
significance of the difference between their ordered levels along with
the corresponding interpretation. This will take us to the concept of
contrasts.

```{r}
options(contrasts = c("contr.treatment", "contr.sdif"))
```

#### Without Confounders

Calculating the $p$-values for the regression coefficients along with
the adjusted $p$-values via the Benjamini-Hochberg procedure with
$\alpha = 0.05$ and Showing the model's summary via the function
`tidy()`.

```{r polr mass }
initial_model <- polr(salary_exp_post_grad ~ mds_self_rated_enjoy, Hess = TRUE, data = salary_df)
summary_initial_model <- tidy(initial_model) %>%
  mutate(p_value = pnorm(abs(statistic), lower.tail = FALSE) * 2) %>%
  mutate(p_value_adj = p.adjust(p_value, method = "BH")) %>%
  mutate_if(is.numeric, round, 2)


print(summary_initial_model)

summary_initial_model[, c(1,7)] %>%
  filter(p_value_adj < 0.05)

```

There is no significant effect of self-rated enjoyment of MDS on salary
expectation after MDS graduation in any of the **successive
differences** between the ordered levels in the factor self-rated
enjoyment of MDS. However, three out of the four intercepts are
significant. This indicates some differences in the ordered categories
in the response variable, salary expectation after MDS graduation, which
would be due to the different counts of each ordered category.

At this point, given that we have not analyzed confounders, we do not
know if this null effect of self-rated enjoyment of MDS on salary
expectation after MDS graduation is due to there truly being no effect
of self-rated enjoyment of MDS, or if a confounder is masking this
effect. To determine this, we need to investigate the effects of all
possible confounders we can.

#### With Confounders

Estimating the regression model called `full_model` of
`salary_exp_post_grad` versus `mds_self_rated_enjoy` **along with the
rest of the survey questions and `year` as STANDALONE confounders**.
Calculating the $p$-values for the regression coefficients along with
the adjusted $p$-values via the Benjamini-Hochberg procedure with
$\alpha = 0.05$ and showing the model's summary via the function tidy().

```{r with-confounders }


full_model <- polr(salary_exp_post_grad ~ ., Hess = TRUE, data = salary_df)
summary_full_model <- tidy(full_model) %>%
  mutate(p_value = pnorm(abs(statistic), lower.tail = FALSE) * 2) %>%
  mutate(p_value_adj = p.adjust(p_value, method = "BH")) %>%
  mutate_if(is.numeric, round, 2)

print(summary_full_model)

summary_full_model[, c(1,7)] %>%
              filter(p_value_adj < 0.05)
```

Now, once we incorporate the confounders, we can see that
`mds_self_rated_enjoy3-2` is significant on the response. Therefore, the
confounders were masking this effect. Moreover, salary before starting
MDS and years of previous work experience are also the only confounders
that bear any statistical evidence that might be associated with salary
expectation after MDS graduation.

In particular, `salary_pre_mds$80,001 to $100,000-$60,000 to $80,000`
and `salary_pre_mdsMore than $120,000-$100,001 to $120,000`, and
`work_exp1 - 4 Years-Less than 1 Year` have adjusted $p$-values of less
than the significance level $\alpha$ (with 5% False discovery rate
adjustment using the Benjamini-Hochberg procedure).

### Is there a statistical difference between the data of both years?

Given the adjusted $p$-value for `year` in the summary of `full_model`,
we can conclude that making a distinction between both years is not
statistically significant.

### Does the model without confounding variables perform better than a model with confounding variables

It is necessary to statistically check whether `full_model` provides a
better data fit than the `initial_model`. Using $\alpha = 0.05$, conduct
the corresponding hypothesis testing. Do not forget to specify the
corresponding hypotheses.

The hypotheses are:

$$H_0:\texttt{initial\_model} \text{ fits the data as good as the } \texttt{full\_model}$$

$$H_a: \text{otherwise}$$

Given the $p$-value obtained in the Likelihood Ratio Test (LRT), we can
conclude that the `full_model` fits the data better than the
`initial_model`.

```{r comparing models}

anova(full_model, initial_model) %>%
   mutate_if(is.numeric, round, 2)

```

### Reduced Model

Using Standalone explanatory varaibles t hatturned out to have at least
one significant regression coefficient from the previous analysis. The
variables are: `salary_pre_mds` and `work_exp`

Regarding our $X$ of interest, we can see that `mds_self_rated_enjoy2-1`
and `mds_self_rated_enjoy3-2` are statistically significant in this
model. Moreover, in terms of the confounders used in this model,
`salary_pre_mds$60,000 to $80,000-Less than $60,000`,
`salary_pre_mds$80,001 to $100,000-$60,000 to $80,000`,
`salary_pre_mdsMore than $120,000-$100,001 to $120,000`, and
`work_exp1 - 4 Years-Less than 1 Year` are significant.

```{r reduced model}

reduced_model <- polr(salary_exp_post_grad ~ mds_self_rated_enjoy +
  salary_pre_mds + work_exp,
Hess = TRUE, data = salary_df
)
summary_reduced_model <- tidy(reduced_model) %>%
  mutate(p_value = pnorm(abs(statistic), lower.tail = FALSE) * 2) %>%
  mutate(p_value_adj = p.adjust(p_value, method = "BH")) %>%
  mutate_if(is.numeric, round, 2)


summary_reduced_model[, c(1, 7)] %>%
  filter(p_value_adj < 0.05)


```

### Pairwise Comparison Between Three Models

Making pairwise comparisons between `initial_model`, `full_model`, and
`reduced_model`.

Based on the testing results, we can conclude that we can use the
`reduced_model`, since it is as good as the `full_model`), which fits
the data better than the `initial_model`.

```{r pairewise-comparison}
# BEGIN SOLUTION

LRT_full_vs_initial <- anova(full_model, initial_model) %>%
  mutate_if(is.numeric, round, 3)

print(LRT_full_vs_initial)

LRT_full_vs_reduced <- anova(full_model, reduced_model) %>%
  mutate_if(is.numeric, round, 3)

print(LRT_full_vs_reduced)

LRT_reduced_initial <- anova(reduced_model, initial_model) %>%
  mutate_if(is.numeric, round, 3)

print(LRT_reduced_initial)

raw_p_values <- c(
  LRT_full_vs_initial$`Pr(Chi)`[2],
  LRT_full_vs_reduced$`Pr(Chi)`[2],
  LRT_reduced_initial$`Pr(Chi)`[2]
)

# Adjusting for multiple comparisons

test_results <- tibble(
  Test = c(
    "LRT Full Model vs. Initial Model",
    "LRT Full Model vs. Reduced Model",
    "LRT Reduce Model vs. Initial Model"
  ),
  raw_p_values
)
test_results <- test_results %>%
  mutate(p_value_adj = p.adjust(raw_p_values, method = "BH")) %>%
  mutate_if(is.numeric, round, 3)

print(test_results)

```

### Conclusion

Firstly, running a baseline model of the salary expectation after
graduation versus the self-rated enjoyment of MDS does not provide any
statistical evidence of an association between both variables. However,
when adding all the potential confounders from the survey, we see that
some of the coefficients for the association of self-rated enjoyment of
MDS are now significant, even after adjusting for multiple comparisons
in the regression coefficients.

Specifically, when we include all the potential confounders, there is
evidence that there is a relationship between self-rated enjoyment of
MDS on salary expectation after MDS graduation when increasing the
self-rated score from 2 to 3. By conducting a Likelihood Ratio Test of
the baseline versus this model, including all confounders, we have
statistical evidence that taking into account these confounders makes
the model fit the data better. Note that the confounders salary before
joining MDS and previous work experience also show statistical evidence
of association with the salary expectation.

To see if we could fit a better model, we refit it using salary before
starting MDS and previous work experience as the only confounders. When
we do this, the model is not significantly different from the model that
includes all confounders. Thus, we choose this simpler model as our
final model. When evaluating this final model's coefficients, there is
specific evidence that there is a relationship between the expected
salary when increasing the self-rated score from 1 to 2 and 2 to 3. This
is true only when controlling for salary's confounding effect before
starting MDS and previous work experience. Hence, it appears that salary
before starting MDS and previous work experience mask the effect of
self-rated enjoyment of MDS on salary expectation after MDS graduation
(i.e., it is only observable when we take salary before starting MDS and
work experience into account in the model). Note that the year where the
survey was conducted does not have a statistical association with the
response.

### Is this relationship causal?

This is an observational study, so we cannot conclusively say, but we
can at least assess the things we would need to assume to draw causal
conclusions from an observational study and temper our causal claims
based on that. The first assumption is that we have identified all
confounders. We cannot prove this entirely. However, we did assess many
confounders, and only two of the ones we assessed were significant. So
we have some slight evidence towards this assumption.

There might be something else that causes the variability observed in
salary expectation after MDS graduation. However, we do not currently
know what it is. If we were interested in further investigating this, it
would be best to perform an additional observational study (this is not
a study we could experiment for). We would test whether another variable
is causal and again try to control for all possible confounders. Given
what we learned about salary before starting MDS and salary expectation
after MDS graduation, along with previous work experience, we would
likely want to include these two variables in our new analysis as
confounders.

### Possible Improvement

Possible improvements to this analysis includes:

-   A larger sample size.
-   Asking for survey respondents to answer salary and year-related
    questions via inputting a number instead of a drop-down menu. This
    would allow the analysis to be done using linear regression (more
    straightforward to interpret and potentially more powerful).
-   Alternatively, perhaps different boundaries could be chosen if they
    really want to stick with ordered categories. We could use what we
    learned from this study to pick better category boundaries for the
    future.
-   Perform a cohort study so that some covariate questions could be
    asked at a more appropriate time (e.g., how confident in your data
    science skill set did you feel when first starting MDS on a scale of
    1 - 4?) and the explanatory variable of interest (perceived
    enjoyment in MDS could be asked after completion of MDS, i.e., after
    the Capstone project experience).
-   The final model assumes that the $(X,Y)$ association is the same
    across strata (i.e., there are no interactions). This assumption
    might not hold. We can add an interaction term between the
    self-rated enjoyment of MDS and salary before starting MDS and
    previous work experience. If we see significant interaction
    coefficients, this would suggest that the $(X,Y)$ association is not
    the same across strata.

### Attribution

The question, data, and analysis that makes up this causal analysis were
derived from a survey and analysis performed by the following past MDS
students:

-   Carrie Cheung.

-   Alex Pak.

-   Talha Siddiqui.

-   Evan Yathon.
